Robust Control of a Fluxonium Qubit

### 2020-11-18 paper/1introduction.tex
full build
brian edits

Analytic and numerical techniques for QOC have been remarkably successful in
designing high fidelity gates (citations needed). In addition to the high
simulated fidelities which can be obtained with these techniques, it is desirable for the
resultant quantum state trajectory to be insensitive to experimental errors
such as parameter drift, noise, and finite control resolution. Quantum state
trajectories that are insensitive to these effects are said to be robust.
In line with this objective, recent work has sought to mitigate decoherence
and errors due to parameter deviations.
Considering the dynamical and geometric phases of quantum state
trajectories has led to analytic methods for mitigating
errors due to parameter deviations and pure dephasing
\cite{han2020experimental, merrill2014progress, xu2020nonadiabatic, zhang2020universal}.
Floquet techniques have been experimentally demonstrated to simultaneously mitigate
decoherence due to longitudinal relaxation and pure dephasing
\cite{huang2020engineering, mundada2020floquet}.
Numerical QOC techniques have been adapted to mitigate longitudinal relaxation
by modeling master equations \cite{rembold2020introduction} and employing
Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.
Additionally, efforts have been made to incorporate experimental feedback
into optimization \cite{huang2020engineering}. Comprehensive tools that interleave characterization
and pulse design are under active development \cite{wittler2020integrated}.

In this work we develop robust trajectory optimization techniques to mitigate
errors arising from longitudinal relaxation, parameter deviations, and 1/$f$ flux noise
on the fluxonium qubit presented in \cite{zhang2020universal}. We perform these optimization using the ALTRO
method \cite{howell2019altro} which combines the iterative, indirect shooting
flavor of methods such as GOAT \cite{machnes2015gradient}, GRAPE
\cite{khaneja2005optimal, leung2017speedup}, and Krotov's \cite{goerz2019krotov}
with the augmented Lagrangian method. The augmented Lagrangian method
allows us to enforce arbitrary, simultaneous constraints and achieve
fast convergence without being restricted to the constraint
manifold. To mitigate errors due to longitudinal
relaxation, we perform time-optimal control and
encode the dependence of longitudinal relaxation
on the controls in an efficient optimization objective that does
not pay the increased computational cost of integrating a master equation.
Additionally, we consider three robust
robust trajectory opitmization techniques to make the
quantum state trajectory robust to parameter deviations.
The first is the sampling method, which has been applied
previously in the context of QOC
\cite{carvalho2020error, reinhold2019controlling, rembold2020introduction}.
The second is the unscented sampling
method, which derives from the unscented transformation used
for non-linear Kalman filtering
\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
We propose a third method, which uses derivative information
of parameter deviations with respect to the quantum state trajectory.
We make this method efficient by employing mixed-mode differentiation.

This paper is organized as follows.
First, we introduce the ALTRO method in the context of QOC.
We describe realistic constraints for the fluxonium and
map them to the ALTRO method. Then, we
outline a method for making the optimization aware of longitudinal
relaxation. Next, we outline three methods for achieving
robustness to static parameter deviations. Finally,
we employ the robust control techniques to mitigate 1/$f$ flux noise.

>>>>

Analytic and numerical techniques for QOC have been remarkably successful in
designing high fidelity gates (citations needed). 
While these methods are able to predict system behavior with high accuracy, they tend 
to be sensitive to experimental errors such as parameter drift, noise, and finite control 
resolution. 
% In addition to the high
% simulated fidelities which can be obtained with these techniques, it is desirable for the
% resultant quantum state trajectory to be insensitive to experimental errors
% such as parameter drift, noise, and finite control resolution. Quantum state
% trajectories that are insensitive to these effects are said to be robust.
% In line with this objective, 
In order to address these issues,
recent work has sought to mitigate decoherence
and errors due to parameter deviations
Considering the dynamical and geometric phases of quantum state
trajectories has led to analytic methods for mitigating
errors due to parameter deviations and pure dephasing
\cite{han2020experimental, merrill2014progress, xu2020nonadiabatic, zhang2020universal}.
Floquet techniques have been experimentally demonstrated to simultaneously mitigate
decoherence due to longitudinal relaxation and pure dephasing
\cite{huang2020engineering, mundada2020floquet}.
Numerical QOC techniques have been adapted to mitigate longitudinal relaxation
by modeling master equations \cite{rembold2020introduction} and employing
Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.
Additionally, efforts have been made to incorporate experimental feedback
into optimization \cite{huang2020engineering}. Comprehensive tools that interleave characterization
and pulse design are under active development \cite{wittler2020integrated}.

In this work we mitigate
errors arising from longitudinal relaxation, parameter deviations, and 1/$f$ flux noise
on the fluxonium qubit \cite{zhang2020universal} by solving a robust trajectory 
optimization problem. 
Leveraging recent advances in trajectory optimization within the field of robotics, we
solve this optimization problem using the ALTRO solver, which uses iterative LQR 
(iLQR)---a differential-dynamic programming (DDP)-based indirect method similar to shooting
methods such as GOAT \cite{machnes2015gradient}, GRAPE
\cite{khaneja2005optimal, leung2017speedup}, and Krotov's \cite{goerz2019krotov}---within 
an augmented Lagrangian framework to handle nonlinear equality and inequality constraints at 
each time step \cite{howell2019altro}. 
% The augmented Lagrangian method
% allows us to enforce arbitrary, simultaneous constraints and achieve
% fast convergence without being restricted to the constraint
% manifold. 
To mitigate errors due to longitudinal
relaxation, we perform time-optimal control and
encode the dependence of longitudinal relaxation
on the controls in an efficient optimization objective that does
not pay the increased computational cost of integrating a master equation.
Additionally, to make the quantum state trajectory robust to parameter deviations,
we consider three different approaches to making the optimized trajectory less 
sensitive to model uncertainty:
% we consider three robust
% robust trajectory opitmization techniques to make the
% quantum state trajectory robust to parameter deviations.
\begin{enumerate}
  \item A sampling method, similar to the work of \cite{carvalho2020error, reinhold2019controlling, rembold2020introduction}.
  \item An unscented sampling method, adapted from the unscented tranform used in the 
  state estimation community \cite{howell2020direct, julier2004unscented, lee2013sigma, manchester2016derivative}.
  \item A derivative-penalization method, which uses efficient mixed-mode differentiation
  to compute derivative information of parameter deviations with respect to the 
  quantum state trajectory.
\end{enumerate}
% The first is the sampling method, which has been applied
% previously in the context of QOC
%%TODO: the citations for unscented sampling
%% and sampling are the same as in section V
% \cite{carvalho2020error, reinhold2019controlling, rembold2020introduction}.
% The second is the unscented sampling
% method, which derives from the unscented transformation used
% for non-linear Kalman filtering
% \cite{howell2020direct, julier2004unscented, lee2013sigma, manchester2016derivative}.
% We propose a third method, which uses derivative information
% of parameter deviations with respect to the quantum state trajectory.
% We make this method efficient by employing mixed-mode differentiation.

This paper is organized as follows. \todo{use section labels?}
First, we introduce the ALTRO method in the context of QOC.
We describe realistic constraints for the fluxonium and
map them to the ALTRO method. Then, we
outline a method for making the optimization aware of longitudinal
relaxation. Next, we outline three methods for achieving
robustness to static parameter deviations. Finally,
we employ the robust control techniques to mitigate 1/$f$ flux noise.

### 2020-11-18 paper/2background.tex
brian edits

The trajectory optimization
literature solves a more general class of non-linear programs that resemble
the QOC problem. The QOC
problem is a specific case of the linear quadratic regulator (LQR).
LQR is so called because the dynamics are linear in the state and
the functional is quadratic in the state. A general LQR cost
function takes the form
\begin{equation}
  J_{\textrm{iLQR}} = \ell_{N}({x}_{N})
  + \sum_{k = 0}^{N - 1} \ell_{k}({x}_{k}, u_{k})
\end{equation}
The subscript $k$ denotes the knot point, $x$ is the augmented state vector,
and $u$ is the augmented control vector. The augmented state is propagated
to the next knot point by the dynamics function $x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
In the case of QOC $\ket{\psi_{k}} \subseteq x_{k}$
and $f$ encodes the TDSE dynamics \eqref{eq:tdse}. The
cost function $\ell_{k}$ includes contributions from the infidelity
$1 - {\lvert \braket{\psi_{f} \lvert \psi_{k}} \rvert}^{2}$, final state deviations
$\tilde{x}^{T}_{k} Q_{k} \tilde{x}_{k}$ where $\tilde{x}_{k} = x_{f} - x_{k}$,
and control regularization $u^{T}_{k} R_{k} u_{k}$.
Placing cost functions at each knot point smoothens the optimization
landscape. In QOC realizing the final state
is the most important objective so the cost function at the final knot point
carries a higher realtive weight $\ell_{N} \sim N \cdot \ell_{k}$.
The LQR formulation restricts the cost functions to operate on the augmented
state and controls at each knot point. The advantage of the LQR formulation
is that there exists a differentiable dynamic programming algorithm to compute the
optimal update to the augmented controls $u_{k}$ which minimizes the cost
function $\ell_{k}$ at each knot point. This algorithm proceeds by deriving a
recurrence relation between knot points $k$ and $k + 1$ for the optimal
feedback law--known as the Ricatti recursion (citations needed). The
iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
and applies the Ricatti recursion to all knot points on multiple
executions.

In order to incorporate constraints we employ
the augmented Lagrangian method. Constraints are contributions
to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
zero or negative when the constraint is satisfied. The AL-iLQR
method associates a penalty multiplier with the constraint
that estimates the constraint's Lagrange multiplier.
The algorithm updates the penalty multiplier between
iLQR executions. In this scheme the functional takes the form
\begin{equation}
\begin{aligned}
    J_{\textrm{AL-iLQR}} &= J_{\textrm{iLQR}}\\
    &+ (\lambda_{N} + \frac{1}{2}I_{\mu_{N}} c_{N}(x_{N}))^{T} c_{N}(x_{N})\\
    &+ \sum^{N - 1}_{k = 0} (\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
\end{aligned}
\end{equation}
Here $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
with $\mu_{k}$ along the diagonal.
$\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
\begin{align}
  \lambda^{i}_{k} &\gets \max(0, \lambda^{i}_{k} + \mu^{i}_{k} c^{i}_{k}(x_{k}^{*}, u_{k}^{*}))\\
  \mu^{i}_{k} &\gets \phi \mu^{i}_{k}
\end{align}
Here $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
$i$ indicates the $i$-th constraint functional,
and $\phi$ is a hyperparameter. An optimal control update, similar to the Ricatti
recursion, can be formed to calculate the optimal control updates for this functional,
see \cite{howell2019altro}.

>>>>

Alternatively, the QOC problem can be formulated as a trajectory optimization problem 
and solved using any of the state-of-the-art solvers developed by the robotics community.
Trajectory optimization problems are typically of the following form: 
\begin{mini}[2]
    {x_{0:N},u_{0:N-1}}{\ell_f(x_N) + \sum_{k=0}^{N-1} \ell(x_k,u_k) }{}{}
    \addConstraint{x_{k+1} = f(x_k,u_k)}
    \addConstraint{g_k(x_k,u_k)}{\leq 0}
    \addConstraint{h_k(x_k,u_k)}{=0}
    \label{opt:discrete_trajopt},
\end{mini}
where $\ell_f$ and $\ell$ are the final and stage cost functions, $x_k \in \R^n$ and
$u_k \in \R^m$ are the state and input control variables, $f(x_k,u_k)$ is the discrete
dynamics function, and $g_k(x_k,u_k)$ and $h_k(x_k,u_k)$ are the inequality
and equality constraints, potentially including initial and final conditions,
at time step $k$.

Many techniques have been proposed for solving \eqref{opt:discrete_trajopt}. Standard 
methods include direction collocation \cite{Hargraves87} and differential-dynamic programming
(DDP) \cite{Mayne1966a}. Recent state-of-the-art solvers, such as ALTRO \cite{howell2019altro},
have combined principles from both of these approaches.

ALTRO uses iterative LQR (iLQR) \cite{Li2004a} as the internal solver of an augmented 
Lagrangian method (ALM). iLQR solves an unconstrained trajectory optimization problem 
using a backward Riccati recursion to derive a closed-loop linear feedback law about the current 
trajectory. By simulating the system forward with the feedback law, the trajectory is 
brought closer to the (locally) optimal trajectory. DDP-based solvers such as iLQR
are popular since they are very computationally efficient, are always dynamically feasible,
and provide a closed-loop control policy about the optimal trajectory for free; however,
standard implementations have no ability to deal with nonlinear equality and inequality 
constraints. ALM handles constraints by successively solving unconstrained minimization 
problems of the form:
\begin{mini}[2]
    {z}{f(z) + \lambda^T c(z) + \half c(z)^T I_\mu c(z)}{}{}
    \label{opt:alm}
\end{mini}
where $f(z)$ is the objective function, $c(z) : \R^n \mapsto \R^p$ is the constraint 
function, $\lambda \in \R^p$ are the Lagrange multipliers, and $I_\mu$ is a diagonal matrix
of penalty weights, whose magnitude depend on whether the constraint is active or inactive.
For ALTRO, $f(z)$ is the objective function of \eqref{opt:discrete_trajopt}, $c(z)$ is the 
concatenation of $g_k$ and $h_k$, and $z$ is the concatenation of the states and controls 
across all time steps. After minimizing \eqref{opt:alm} using iLQR, the penalty terms and 
Lagrange multipliers are updated, and the process repeats until convergence.

ALM converges superlinearly but tends to exhibit slow constraint convergence near the optimal 
solution due to poor numerical conditioning. To address this shortcoming, ALTRO provides a 
solution-polishing phase that takes 1-2 Newton steps on the active constraint set to provide 
machine-precision constraint satisfaction. For more information on the details of the ALTRO
solver, see \cite{howell2019altro,Jackson2020altroc}.

% For the QOC problem, the state $x$ consists of \todo{fill this out} while the control input
% $u$ includes the flux drive amplitude. The cost functions $\ell_N$
% The trajectory optimization
% literature, driven primarily by the robotics community,
% solves a more general class of non-linear programs that resemble
% the QOC problem. The QOC
% problem is a specific case of the linear quadratic regulator (LQR).
% LQR is so called because the dynamics are linear in the state and
% the functional is quadratic in the state. A general LQR cost
% function takes the form
% \begin{equation}
%   J_{\textrm{iLQR}} = \ell_{N}({x}_{N})
%   + \sum_{k = 0}^{N - 1} \ell_{k}({x}_{k}, u_{k})
% \end{equation}
% The subscript $k$ denotes the knot point, $x$ is the augmented state vector,
% and $u$ is the augmented control vector. The augmented state is propagated
% to the next knot point by the dynamics function $x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
% In the case of QOC $\ket{\psi_{k}} \subseteq x_{k}$
% and $f$ encodes the TDSE dynamics \eqref{eq:tdse}. The
% cost function $\ell_{k}$ includes contributions from the infidelity
% $1 - {\lvert \braket{\psi_{f} \lvert \psi_{k}} \rvert}^{2}$, final state deviations
% $\tilde{x}^{T}_{k} Q_{k} \tilde{x}_{k}$ where $\tilde{x}_{k} = x_{f} - x_{k}$,
% and control regularization $u^{T}_{k} R_{k} u_{k}$.
% Placing cost functions at each knot point smoothens the optimization
% landscape. In QOC realizing the final state
% is the most important objective so the cost function at the final knot point
% carries a higher realtive weight $\ell_{N} \sim N \cdot \ell_{k}$.
% The LQR formulation restricts the cost functions to operate on the augmented
% state and controls at each knot point. The advantage of the LQR formulation
% is that there exists a differentiable dynamic programming algorithm to compute the
% optimal update to the augmented controls $u_{k}$ which minimizes the cost
% function $\ell_{k}$ at each knot point. This algorithm proceeds by deriving a
% recurrence relation between knot points $k$ and $k + 1$ for the optimal
% feedback law--known as the Ricatti recursion (citations needed). The
% iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
% and applies the Ricatti recursion to all knot points on multiple
% executions.

% In order to incorporate constraints we employ
% the augmented Lagrangian method. Constraints are contributions
% to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
% zero or negative when the constraint is satisfied. The AL-iLQR
% method associates a penalty multiplier with the constraint
% that estimates the constraint's Lagrange multiplier.
% The algorithm updates the penalty multiplier between
% iLQR executions. In this scheme the functional takes the form
% \begin{equation}
% \begin{aligned}
%     J_{\textrm{AL-iLQR}} &= J_{\textrm{iLQR}}\\
%     &+ (\lambda_{N} + \frac{1}{2}I_{\mu_{N}} c_{N}(x_{N}))^{T} c_{N}(x_{N})\\
%     &+ \sum^{N - 1}_{k = 0} (\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
% \end{aligned}
% \end{equation}
% Here $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
% with $\mu_{k}$ along the diagonal.
% $\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
% \begin{align}
%   \lambda^{i}_{k} &\gets \max(0, \lambda^{i}_{k} + \mu^{i}_{k} c^{i}_{k}(x_{k}^{*}, u_{k}^{*}))\\
%   \mu^{i}_{k} &\gets \phi \mu^{i}_{k}
% \end{align}
% Here $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
% $i$ indicates the $i$-th constraint functional,
% and $\phi$ is a hyperparameter. An optimal control update, similar to the Ricatti
% recursion, can be formed to calculate the optimal control updates for this functional,
% see \cite{howell2019altro}.

### 2020-11-18 paper/3fluxonium.tex
brian edits

In addition to imposing optimization constraints that
reflect physical limitations of the apparatus, we impose
constraints that improve the experimental realization of the control pulse.
To ensures gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions,
we require $a(t = 0) = a(t = t_{N}) = 0$.
Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines,
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
to ensure the two-level approximation \eqref{eq:hamiltonian}
remains valid. Additionally, we require that each gate achieves
the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
In addition to these constraints we penalize the norm
of the first and second derivatives of the flux amplitude to
ensure its smoothness, and in doing so mitigate AWG ringing.
The optimization is performed over the second derivative of the flux amplitude
$\frac{d^{2}}{dt^{2}} a$ which is contained in the
augmented control vector. The first derivative
$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
flux amplitude terms
are contained in the augmented state vector. They are obtained from
the second derivative of the flux amplitude by
integration in the dynamics function.
Both the zero net flux and target quantum state constraint
are then handled by ensuring the target augmented state is
reached $x_{N} = x_{f}$, the constraint functional takes the form
$c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
The equality and inequality constraints on $a$ are handled
with a bound constraint which takes the form
$c_{k} = (x_{k} - b)^{T}(x_{k} - b)$ if $\lvert x_{k} \rvert \ge b$
and $0$ otherwise.
In addition the norm of the quantum state
is constrained to ensure discretization error is not
exploited $c_{k} = {\lvert \braket{\psi_{k} \lvert \psi_{k}} \rvert}^{2} - 1$.

@@

>>>>

\todo{I highly suggest writing out the optimization problem explicitly here. I've started it 
for you below. You can reference the individual lines as you explain your problem setup.
Also explicitly spell out what your state and control vectors are.}
{\color{red}
\begin{mini!}[2]
    {x_{0:N}, u_{0:N\text{-}1}}{\sum_{k=1}^N \norm{x_k\text{-}x_f}_{Q_k} + \sum_{k=1}^{N-1} \norm{u_k}_{R_k}}{}{}
    \addConstraint{x_{k+1}}{= f(x_k, u_k)}  \label{eq:dyn_con}
    \addConstraint{x_1 }{= x_\text{init}} \label{eq:init_con}
    \addConstraint{x_N}{= x_f} \label{eq:term_con}
\end{mini!}
}

In addition to imposing optimization constraints that
reflect physical limitations of the apparatus, we impose
constraints that improve the experimental realization of the control pulse.
To ensures gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions,
we require $a(t = 0) = a(t = t_{N}) = 0$.
Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines,
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
to ensure the two-level approximation \eqref{eq:hamiltonian}
remains valid. Additionally, we require that each gate achieves
the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
In addition to these constraints we penalize the norm
of the first and second derivatives of the flux amplitude to
ensure its smoothness, and in doing so mitigate AWG ringing.
The optimization is performed over the second derivative of the flux amplitude
$\pdv*[2]{a}{t}$ which is contained in the
augmented control vector. The first derivative
$\pdv*{a}{t}$, proportional $a$, and integral $\int a$
flux amplitude terms
are contained in the augmented state vector. They are obtained from
the second derivative of the flux amplitude by
integration in the dynamics function.
Both the zero net flux and target quantum state constraint
are then handled by ensuring the target augmented state is
reached $x_{N} = x_{f}$, the constraint functional 
\todo{not sure what the convention is in your community, but for us a functional is a 
function of functions (or trajectories). This would just be a function. Also, shouldn't 
this just be $x_N = x_f$ and $\abs{x_k} \le b$? Why are you squaring the functions?}
 takes the form
$c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
The equality and inequality constraints on $a$ are handled
with a bound constraint which takes the form
$c_{k} = (x_{k} - b)^{T}(x_{k} - b)$ if $\lvert x_{k} \rvert \ge b$
and $0$ otherwise.
In addition the norm of the quantum state
is constrained to ensure discretization error is not
exploited $c_{k} = {\lvert \braket{\psi_{k} \lvert \psi_{k}} \rvert}^{2} - 1$.
\todo{is this last constraint supposed to be an inequality constraint?}

@@

\todo{maybe reference the next Sections as you mention these points?}

### 2020-11-18 paper/4longitude.tex
paper updates
removed old dirs
full build
brian edits

%% F1
\begin{figure*}[ht]
  \label{fig:longitude}
  \begin{subfigure}{.315\textwidth}
    \includegraphics[width=\linewidth]{assets/f1a.png}
  \end{subfigure}\hfill
  \begin{subfigure}{.23\textwidth}
    \includegraphics[width=\linewidth]{assets/f1b.png}
  \end{subfigure}\hfill
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f1c.png}
  \end{subfigure}
  \caption{
    (a) Numerically optimized gates (red) and analytically optimized gates (blue).
    (b) $T_{1}$ interpolation function used in optimization. Markers
    denote the time-averaged, absolute amplitude of each gate.
    (c) Lindblad master equation simulation with $T_{1}$ dissipation
    for successive gate applications. The reported gate error is the cumulative
    gate error after each gate application.
    The numerically optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
    in the figure.
  }
\end{figure*}

%% S4
\section{Longitudinal Relaxation Awareness}
The longitudinal relaxation time $T_{1}$ varies with
control parameters in a range of superconducting circuit platforms.
It is advantageous to tune the controls to extend the longitudinal
relaxation time.
Computing the gate error due to longitudinal relaxation
requires propagating density matrices of size $n \times n$ under master equation
dynamics, rather than state vectors of size $n$ under the TDSE dynamics.
We avoid this increase in computational complexity by
penalizing the integrated rate of longitudinal relaxation,
i.e. the probability of longitudinal relaxation.
Using this probability as proxy for the gate error incurred
is reasonable because losses due to longitudinal relaxation
increase monotonically in time.
This technique can be extended to
error channels which share the monotonically increasing property.
Additionally, for a constant $T_{1}$ time, a shorter gate duration
would favor a lower longitudinal relaxation probability. We allow
the optimizer to tune the gate duration in order to minimize the
longitudinal relaxation probability. Our scheme for time-optimal
control is applicable to any time-optimal problem, not only
the one we study here.

@@

%% F2
\begin{figure*}[ht]
  \label{fig:static}
  \begin{subfigure}{.315\textwidth}
    \includegraphics[width=\linewidth]{assets/f2a.png}
  \end{subfigure}\hfill
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f2c.png}
  \end{subfigure}\hfill
  \begin{subfigure}{.23\textwidth}
    \includegraphics[width=\linewidth]{assets/f2b.png}
  \end{subfigure}

  \caption{
    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
    analytic, sampling, unscented sampling, and the 1\textsuperscript{st}-
    and 2\textsuperscript{nd}-order derivative methods. The gates
    shown for the numerical methods are the solutions at twice the analytic
    gate time.
    (b) Single gate error as
    a function of the gate duration at a one-percent
    detuning from the nominal qubit frequency for all methods. Missing
    data points represent solutions with a gate error above $5 \cdot 10^{-5}$.
    (c) Single gate error as a function of the detuning from the nominal
    qubit frequency. The solutions for the analytic and
    1\textsuperscript{st}-order derivative methods are shown at multiples
    of the analytic gate time. The performance of the two methods is
    indistinguishable at the analytic gate time $18$ns.
  }
\end{figure*}

We compare the numerical method we have developed to the analytic gates
on the task of achieving low gate errors in the presence of longitudinal relaxation
for the $Z/2$, $Y/2$, and $X/2$ gates.
The numerically optimized gates converge on similar solutions, a periodic
waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure \ref{fig:longitude}.
They extend their gate times
beyond their analytic counterparts, trading longer gate times for access
to higher amplitudes and therefore higher $T_{1}$ times. All numerical gates reduce
their single gate errors by a factor of $5$ over
their analytic counterparts which is commensurate to their
probability of longitudinal relaxation reductions, see Appendix \ref{appendix:longitude}.
The gate error reported in this text is the infidelity
of the evolved state and the target state averaged over 1000 pseudo-randomly
generated initial states. The numerical $Z/2$ and $Y/2$ gates perform simlarly in
the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The numerical $X/2$ gate
achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
Both the analytic and numerical gates attain single gate errors sufficient for
quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
The low gate errors achieved by the numerical gates for extend computations
are critical for noisy, intermediate-scale quantum (NISQ) applications.
These improvements are significant for the realistic constraints we have imposed
on the gates, and do not represent a fundamental limit to the optimization methods we have
employed.

>>>>

\begin{figure*}[ht]

  \begin{subfigure}{.315\textwidth}
    \includegraphics[width=\linewidth]{assets/f1a.png}
  \end{subfigure}\hfill
  \begin{subfigure}{.23\textwidth}
    \includegraphics[width=\linewidth]{assets/f1b.png}
  \end{subfigure}\hfill
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f1c.png}
  \end{subfigure}
  \caption{
    (a) Numerically optimized gates (red) and analytically optimized gates (blue).
    (b) $T_{1}$ interpolation function used in optimization. Markers
    denote the time-averaged, absolute amplitude of each gate.
    (c) Lindblad master equation simulation with $T_{1}$ dissipation
    for successive gate applications. The cumulative
    gate error is computed after each gate application.
    The numerically optimized $Z/2$ and $Y/2$ gate errors are indistinguishable
    in the figure.
  }
  \label{fig:longitude}
\end{figure*}

\section{Longitudinal Relaxation Awareness}
The longitudinal relaxation time $T_{1}$ varies with
control parameters in a range of superconducting circuit platforms.
It is advantageous to tune the controls to extend the longitudinal
relaxation time.
Computing the gate error due to longitudinal relaxation
requires propagating density matrices of size $n \times n$ under master equation
dynamics, rather than state vectors of size $n$ under the TDSE dynamics.
We avoid this increase in computational complexity by
penalizing the integrated rate of longitudinal relaxation,
i.e. the probability of longitudinal relaxation.
Using this probability as proxy for the gate error incurred
is reasonable because losses due to longitudinal relaxation
increase monotonically in time.
This technique can be extended to
error channels which share the monotonically increasing property.
Additionally, for a constant $T_{1}$ time, a shorter gate duration
would favor a lower longitudinal relaxation probability. We allow
the optimizer to tune the gate duration in order to minimize the
longitudinal relaxation probability. Our scheme for time-optimal
control is applicable to any time-optimal problem, not only
the one we study here.

\todo{again, I'd recommend writing out either the entire optimization problem or at least
point to the parts of the optimization problem defined in the previous section you had to 
change. Just make it a little more concrete.}

@@

%% F2
\begin{figure*}[ht]
  \begin{subfigure}{.315\textwidth}
    \includegraphics[width=\linewidth]{assets/f2a.png}
  \end{subfigure}\hfill
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f2c.png}
  \end{subfigure}\hfill
  \begin{subfigure}{.23\textwidth}
    \includegraphics[width=\linewidth]{assets/f2b.png}
  \end{subfigure}
  \caption{
    (a) $Z/2$ gates robust to qubit frequency detunings constructed with the
    analytic, sampling, unscented sampling, and the 1\textsuperscript{st}-
    and 2\textsuperscript{nd}-order derivative methods. The gates
    shown for the numerical methods are the solutions at twice the analytic
    gate time.
    (b) Single gate error as
    a function of the gate duration at a one-percent
    detuning from the nominal qubit frequency for all methods. Missing
    data points represent solutions with a gate error above $5 \cdot 10^{-5}$.
    (c) Single gate error as a function of the detuning from the nominal
    qubit frequency. The solutions for the analytic and
    1\textsuperscript{st}-order derivative methods are shown at multiples
    of the analytic gate time. The performance of the two methods is
    indistinguishable at the analytic gate time $18$ns.
  }
  \label{fig:static}
\end{figure*}

We compare the numerical method we have developed to the analytic gates
on the task of achieving low gate errors in the presence of longitudinal relaxation
for the $Z/2$, $Y/2$, and $X/2$ gates.
The numerically optimized gates converge on similar solutions, a periodic
waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure \ref{fig:longitude}.
They extend their gate times
beyond their analytic counterparts, trading longer gate times for access
to higher amplitudes and therefore higher $T_{1}$ times. All numerical gates reduce
their single gate errors by a factor of $5$ over
their analytic counterparts which is commensurate to their
probability of longitudinal relaxation reductions, see Appendix \ref{appendix:longitude}.
The gate error reported in this text is the infidelity
of the evolved state and the target state averaged over 1000 pseudo-randomly
generated initial states. The numerical $Z/2$ and $Y/2$ gates perform simlarly in
the concatenated gate application comparision, suppressing accumulated gate errors to $8 \cdot 10^{-3}$
over $2000$ gate applications $\sim 40 \mu\textrm{s}$. The numerical $X/2$ gate
achieves an accumulated gate error of $1.7 \cdot 10^{-2}$ over $2000$ gate applications $\sim 124 \mu\textrm{s}$.
Both the analytic and numerical gates attain single gate errors sufficient for
quantum error correction $< 10^{-4}$, requisite for fault-tolerant quantum computing.
The low gate errors achieved by the numerical gates \todo{typo / awkward sentence} 
for extend computations
are critical for noisy, intermediate-scale quantum (NISQ) applications.
These improvements are significant for the realistic constraints we have imposed
on the gates, and do not represent a fundamental limit to the optimization methods we have
employed.
